---
title: "Data Analysis of an Insurance Company Result"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```







```{r}
library(tidyverse)

claims_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/claims_df.rds'))

```


### Raw Data

```{r}

claims_df

```
<br>
<br>


# Exploratory Data Analysis

## Question 1

Which state is responsible for the majority of the profits?



**Answer**:
California has the highest monetary profit at $1898706 which constitues 32.95% of the overall profit. Although the average profit is comparitively lower, it still rakes in the highest profit margin.
The total number of claims in California is 5185.


States like Washington and Nevada has the least number of claims.


To add additional R code chunks for your work, select `Insert` then `R` from the top of this notebook file.

```{r}
# Question 01

library(ggplot2)
library(tidyr)
library(dplyr)
library(tidyverse)
library(skimr)
library(usmap)
claims_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/claims_df.rds'))
claims_df = claims_df %>% mutate(revenue = monthly_premium * months_policy_active)
claims_df_1 = claims_df %>% mutate(state = claims_df$customer_state)
profit = sum(claims_df$customer_lifetime_value)
profit
q1 = claims_df_1 %>% group_by(state) %>% summarize(
  total_revenue = sum(revenue),
  count = n(),
  total_claims = sum(total_claims),
  sum_amount_claimed = sum(total_claims_amount),
  totalprofit = sum(customer_lifetime_value),
  avg_profit = mean(customer_lifetime_value),
  percentage_profit = (totalprofit/5761975) *100)
q1



plot_usmap(data = q1,values = "totalprofit",include = c("CA","WA","OR","NV","AZ"), labels = TRUE) + 
  scale_fill_gradient(name = "Profit", low = "pink", high =  "red", na.value = "grey50") +labs(title = "Statewise customer value")

```



## Question 2

How is Monthly premiun paid affect company profits?

**Answer**:

Customers paying higher monthly premium( $100 - $150), although  1835 customers pay in this range. They contribute to an total profit of $3219546
However when you compare this to the most widely paid montly premium in the range of ($50 - $100) by 4052 customers , the profit obtained is $530130. 


```{r}
ggplot(claims_df,aes(x = monthly_premium)) + geom_histogram(fill = "#006EA1", color = "white" , bins = 30) +
  labs(title =" Monthly premium paid by Customers", x = "Monthly premium amount")
claims_df_updates_3 = claims_df %>% mutate(monthly_premium_range = cut_width(monthly_premium, width = 50, boundary = 0))
#summary stats


claims_df_updates_3 %>% group_by(monthly_premium_range) %>% summarize(
  count = n(),
  avg_monthly_premium = mean(monthly_premium),
  avg_cust_value = mean(customer_lifetime_value),
  total_profit = sum(customer_lifetime_value)
)




```



## Question 3

Is the profit of the company getting affected by the type of coverage chosen by the customers along with the monthly premium paid?

**Answer**:

Company faces loss of around 47% in the Basic coverage opted by the 3815 customers paying an average of $82 per month.Major loss is incurred by the customer choosing basic coverage and paying monthly premium in the range of $61 to $75.

Executive coverage also faces loss of 18% opted by 1858 customer whose average monthly membership is $104.

Premium coverage faces the least loss of 10% opted by 576 customers.






```{r}

ggplot(claims_df, aes(x = monthly_premium, y = customer_lifetime_value)) + 
  geom_point(color = "green") +facet_wrap(~coverage ,nrow=3) +labs(title = "Relationship between Monthly premium,Customer_lifetime_value and the coverage ")

#Summary needs to be done
claims_df %>% group_by(coverage) %>% summarize(
  avg_monthly_premium = mean(monthly_premium),
  count= n(),
  avg_loss_cust = mean(customer_lifetime_value<0)
)




```




## Question 4

How is company's profit affected by the different sales channel?

**Answer**:

Out of 4 sales channel, 2359 customers who took up coverage through third party agents have fetched highest profit of $2288145. 
Customers joining the company through Branch, call center and web brings around same average profit.
Web sales channel has  attracted 901 customers which is the least.

```{r}
hw <- theme_gray()+ theme(
  plot.title=element_text(hjust=0.5),
  plot.subtitle=element_text(hjust=0.5),
  plot.caption=element_text(hjust=-.5),
  # strip.text.y = element_blank(),
  strip.background=element_rect(fill=rgb(.9,.95,1),
                                colour=gray(.5), size=.2),
  panel.border=element_rect(fill=FALSE,colour=gray(.70)),
  panel.grid.minor.y = element_blank(),
  panel.grid.minor.x = element_blank(),
  panel.spacing.x = unit(0.10,"cm"),
  panel.spacing.y = unit(0.05,"cm"),
  # axis.ticks.y= element_blank()
  axis.ticks=element_blank(),
  axis.text=element_text(colour="black"),
  axis.text.y=element_text(margin=margin(0,3,0,3)),
  axis.text.x=element_text(margin=margin(-1,0,3,0),angle=90)
)



ggplot(claims_df, aes(x = customer_lifetime_value)) +
  geom_histogram(fill = "#006EA1", color = "white" , bins = 20) + 
  facet_grid(~sales_channel) +hw
#summary stats needs to be made

claims_df %>% group_by(sales_channel) %>% summarize(
  count= n(),
  total_profit = sum(customer_lifetime_value),
  avg_profit = mean(customer_lifetime_value)
  
)



```




## Question 5

Which residence type has the more more of claims?


**Answer**:

Sub-Urban has the highest number of 2 times claim and 3 times claim accounting for 2193 and 1165 customers respectively.

Second highest claims are seen in urban areas where the number of 2 times claimed account for 885  customers.

```{r}


ggplot(claims_df, aes( x = residence_type, fill = residence_type)) + geom_bar(stat = "count" ) +
  facet_wrap(~total_claims) +hw +labs(title = "Residence type affecting Number of claims")

# summary needs to be done

claims_df %>% group_by(residence_type, total_claims) %>% summarize(
  count = n(),
)





```


## Question 6

Which vehicle class fetches higher profit?

**Answer**:

Luxury Car has not brought any loss to the company from all the three coverage types. Its average loss is 0%.
2-Door cars and 4-Door cars bring the highest loss having  1292 and 3124 claims respectively.

Customer with 2-Door basic coverage and customers with 4-Door basic coverage bring the highest loss to the company.

Average profit for 2-door car is $269 which is the least among other vehicle class.


```{r}
ggplot(claims_df,aes(x = vehicle_class, y = customer_lifetime_value, fill = coverage)) + geom_boxplot() +
  labs(title = "Coverage and Vehicle class affecting profits", x = "Vehicle Class", y = "Profit" ) +hw

claims_df = claims_df %>% mutate(revenue = monthly_premium * months_policy_active)
# profit of the company increases with the luxury SUV
#2 door cars and 4 door cars bring loss


claims_df %>% group_by(vehicle_class) %>% summarize(
  number_of_claims = n(),
  avg_profit = mean(customer_lifetime_value),
  avg_revenue = mean(revenue),
  loss_occured = any(customer_lifetime_value<0),
  avg_loses = mean(customer_lifetime_value < 0)
)






```


## Question 7

Which policies offers higher profit to the company?


**Answer**:

4658 Customers have opted for personal policy which fetched a profit of $4302430. 
Special policies are least bought by the customers whose average profit($745.9582) is less compared to other policies.


```{r}




q4 = claims_df %>% group_by(policy) %>% summarize(
  avg_profit = mean(customer_lifetime_value),
  count = n(),
  total_profit = sum(customer_lifetime_value),
  med_profot = median(customer_lifetime_value)
)
q4

pie(q4$avg_profit , labels = c("Personal","Corporate","Special") , edges=10 , main ="Profit through different Policies"
    ) 





```



## Question 8

What is the relationship between the coverage plan chosen by the customer and their income?

**Answer**:

Customers with income in the range of $20000 to $40000 choose Basic coverage over other plans.


```{r}

claims_df_updates_q8 = claims_df %>% mutate(income_range = cut_width(income, width = 20000, boundary = 0))


ggplot(claims_df_updates_q8, aes( x = income_range )) +
  geom_bar(stat = "count", color = "blue") +hw +labs(title ="Income of the customer and the coverage opted", x = "IncomeRange",y = "count") +facet_wrap(~coverage)



```















